### 2014-06-26 1_introduction/introduction.tex
auto

Thus a significant portion of this research effort focuses on developing a computational framework for simulation, evaluation, processing and classification of target polarimetric signatures.
The benefits of such a framework are manyfold.
Firstly, simulation allows one to quickly carry out experiments on small synthesized and well-designed scenarios, without the need for highly expensive satellite imagery.
Secondly, simulation provides ground-truth, without which quantitative evaluation of estimators would have been impossible.
Last but certainly not least, such a framework allow one to apply estimators repeatedly, and results from different estimators can be statistically and reliably compared. For example, this would allow different authors to compare results quantitatively.

@@

\begin{table}
\centering
\begin{tabular}{|c|c|c|}
\hline
Radar Platform & Polarization Mode & Data Capture Time \\
\hline
RadarSat 2 & Full Quad Pol & 29 Apr 2009 \\
RadarSat 2 & Full Quad Pol & 05 Apr 2009 \\
RadarSat 2 & Full Quad Pol & 12 Mar 2009 \\
RadarSat 2 & Full Quad Pol & 23 Jan 2009 \\
RadarSat 2 & Full Quad Pol & 30 Dec 2008 \\
RadarSat 2 & Full Quad Pol & 06 Dec 2008 \\
RadarSat 2 & Full Quad Pol & 19 Oct 2008 \\
RadarSat 2 & Full Quad Pol & 12 Nov 2008 \\
RadarSat 2 & Full Quad Pol & 25 Sep 2008 \\
TerraSar-X & Dual Pol (HH-VV) & 18 Apr 2008 \\
NASA-JPL AirSar & Full Quad Pol & 2001 \\
\hline
\end{tabular}
\caption{Collected imagery data}
\label{tbl:collected_data}
\end{table}

@@

\section{Research Contributions}
Interestingly, the dual-problems stated above are found to have a single solution.
They are the derivation of scalar, representative and homoskedastic statistical models for the multidimensional, inter-correlated and heteroskedastic POLSAR data.
%  which is

@@

% How do you know if the problem is solved?   State your research objectives
Specifically, the following results are to be obtained:
\begin{enumerate}
\item                Scalar and representative statistical models
                for the multi-dimensional and inter-correlated POLSAR
                data are to be derived.
                \begin{enumerate}
                \item \label{itm:polsar_model} A scalar statistical model for the
                POLSAR observable, being the determinant of the
                covariance matrix, is to be derived.
                \item \label{itm:polsar_discrimination_measure} The statistical models for different discrimination measures which are based on
                the observable are to be established together with its consistent property. %being modelled 
                \item \label{itm:polsar_include_sar} The representation power of these observables is to be illustrated by showing
                that when the multi-channel POLSAR data is collapsed
                into the traditional one-channel SAR data, this
                determinant is transformed into the representative SAR
                intensity.
                \item \label{itm:polsar_validation} The model os to be validated against practical real-life
                captured POLSAR data.
                \end{enumerate}
\item                Homoskedastic statistical models for the
                heteroskedastic (POL)SAR data are to be derived. 
                \begin{enumerate}
                \item \label{itm:log_transform} The fact that (POL)SAR data are heteroskedastic in
                the original domain and logarithmic transformation
                can convert this into a homoskedastic model are to be shown
                \item \label{itm:sar_measures_distance} Homoskedastic measures of distance for
                the multi dimensional POLSAR as well as for the one
                dimensional SAR data are to be derived.
                \item \label{itm:sar_validation} These statistical models are to be validated against both
                practical SAR and POLSAR data
                \end{enumerate}
\item The beneficial applications of these
                statistical models are to be demonstrated. Specifically:
                \begin{enumerate}
                \item \label{itm:filter_vicious_circle} To show how the consistent sense of variance
                helps break the vicious cycle in the problem of SAR
                speckle filtering: a new statistical test for
                homogeneity and a new speckle filter is derived.
                \item \label{itm:log_mse} To show how the consistent sense of subtractive
                distance helps in the problem of evaluating (POL)SAR
                speckle filters: the log-MSE is proposed to combine the
                evaluation of radiometric preservation and speckle
                suppression power.
                \item \label{itm:extend_sar_to_polsar} An example where an existing SAR
                processing technique is extensible towards POLSAR data is to be demonstrated
                \end{enumerate}
\end{enumerate}

@@

 The
                beneficial applications of these models is presented in
                chapter \ref{chap:applications}, where the property of a consistent sense of variance
                 helps to break the vicious cycle in SAR
                speckle filtering.
It
                also shows how MSE can again be applicable in the
                additive and homoskedastic domain to evaluate SAR
                speckle filters.
An
                example of extending existing techniques from SAR to
                POLSAR is also demonstrated in this chapter where the
                MSE evaluation criteria is also shown capable of
                evaluating POLSAR speckle filters.
(i.e. see Research Contribution \ref{itm:filter_vicious_circle} to \ref{itm:extend_sar_to_polsar})
Finally, the contribution of the thesis with respect to the research objectives and research impact is discussed in the concluding chapter \ref{chap:conclusions}.

>>>>

Thus a significant portion of this research effort focuses on developing a computational framework for simulation, evaluation, processing and classification of target polarimetric signatures.
The benefits of such a framework are manyfold.
Firstly, simulation allows one to quickly carry out experiments on small synthesized and well-designed scenarios, without the need for highly expensive satellite imagery.
Secondly, simulation provides ground-truth, without which quantitative evaluation of estimators would have been impossible.
Thirdly, such a framework allows one to apply estimators repeatedly, and results from different estimators can be statistically and reliably compared. For example, this would allow different authors to compare results quantitatively.

@@

\begin{table}
\centering
\caption{Collected imagery data}
\label{tbl:collected_data}
\begin{tabular}{|c|c|c|}
\hline
Radar Platform & Polarization Mode & Data Capture Time \\
\hline
RadarSat 2 & Full Quad Pol & 29 Apr 2009 \\
RadarSat 2 & Full Quad Pol & 05 Apr 2009 \\
RadarSat 2 & Full Quad Pol & 12 Mar 2009 \\
RadarSat 2 & Full Quad Pol & 23 Jan 2009 \\
RadarSat 2 & Full Quad Pol & 30 Dec 2008 \\
RadarSat 2 & Full Quad Pol & 06 Dec 2008 \\
RadarSat 2 & Full Quad Pol & 19 Oct 2008 \\
RadarSat 2 & Full Quad Pol & 12 Nov 2008 \\
RadarSat 2 & Full Quad Pol & 25 Sep 2008 \\
TerraSar-X & Dual Pol (HH-VV) & 18 Apr 2008 \\
NASA-JPL AirSar & Full Quad Pol & 2001 \\
\hline
\end{tabular}
\end{table}

@@

\section{Research Contributions}
Interestingly, the dual-problems stated above are found to have a single solution.
They are the derivation of scalar, representative and homoskedastic statistical models for the multi-dimensional, inter-correlated and heteroskedastic POLSAR data.
%  which is

@@

% How do you know if the problem is solved?   State your research objectives
Specifically, the following results are to be obtained:
\begin{enumerate}
\item                Scalar and representative statistical models
                for the multi-dimensional and inter-correlated POLSAR
                data are to be derived.
                \begin{enumerate}
                \item \label{itm:polsar_model} A scalar statistical model for the
                POLSAR observable, being the determinant of the
                covariance matrix, is to be derived.
                \item \label{itm:polsar_discrimination_measure} The statistical models for different discrimination measures which are based on
                the observable are to be established together with its consistent property. %being modelled 
                \item \label{itm:polsar_include_sar} The representation power of these observables is to be illustrated by showing
                that when the multi-channel POLSAR data is collapsed
                into the traditional one-channel SAR data, this
                determinant is transformed into the representative SAR
                intensity.
                \item \label{itm:polsar_validation} The model is to be validated against practical real-life
                captured POLSAR data.
                \end{enumerate}
\item                Homoskedastic statistical models for the
                heteroskedastic (POL)SAR data are to be derived. 
                \begin{enumerate}
                \item \label{itm:log_transform} The fact that (POL)SAR data are heteroskedastic in
                the original domain and logarithmic transformation
                can convert this into a homoskedastic model are to be shown
                \item \label{itm:sar_measures_distance} Homoskedastic measures of distance for
                the multi dimensional POLSAR as well as for the one
                dimensional SAR data are to be derived.
                \item \label{itm:sar_validation} These statistical models are to be validated against both
                practical SAR and POLSAR data
                \end{enumerate}
\item The beneficial applications of these
                statistical models are to be demonstrated. Specifically:
                \begin{enumerate}
                \item \label{itm:filter_vicious_circle} To show how the consistent sense of variance
                helps break the vicious cycle in the problem of SAR
                speckle filtering: a new statistical test for
                homogeneity and a new speckle filter is derived.
                \item \label{itm:log_mse} To show how the consistent sense of subtractive
                distance helps in the problem of evaluating (POL)SAR
                speckle filters: the log-MSE is proposed to combine the
                evaluation of radiometric preservation and speckle
                suppression power.
                \item \label{itm:extend_sar_to_polsar} An example where an existing SAR
                processing technique is extensible towards POLSAR data is to be demonstrated
                \end{enumerate}
\end{enumerate}

@@

 The
                beneficial applications of these models is presented in
                Chapter \ref{chap:applications}, where the property of a consistent sense of variance
                 helps to break the vicious cycle in SAR
                speckle filtering.
It
                also shows how MSE can again be applicable in the
                additive and homoskedastic domain to evaluate SAR
                speckle filters.
An
                example of extending existing techniques from SAR to
                POLSAR is also demonstrated in this chapter where the
                MSE evaluation criteria is also shown capable of
                evaluating POLSAR speckle filters.
(i.e. see Research Contribution \ref{itm:filter_vicious_circle} to \ref{itm:extend_sar_to_polsar})
Finally, the contribution of the thesis with respect to the research objectives and research impact is discussed in the concluding Chapter \ref{chap:conclusions}.

### 2014-06-26 2/literature_review.tex
auto

The stochastic nature of SAR data arises due to the process of SAR image formation. Specifically, since one SAR resolution cell is very big in comparison to the radar wavelength, it normally consists of many elementary scatterers. The data point of each SAR pixel then arises from the combined interference effect of the backscattered waves from these different scatterers. This effect causes natural variation in SAR intensity image even when the underlying scene is highly homogeneous.

SAR speckle phenomena are explained as the interference of many coherent but dephased backscattering components;
  each reflecting from different and distributed elementary scatterers \cite{Oliver_ProcIEEE_1963, Leith_ProcIEEE_1971}. 
The random nature of the SAR data arises
  due to the unknown and varying location, height and thus distance of each elementary scatterer;
  which, in turn, generates random phases in their back-scattering responses.
%Assuming the number of these elementary back scatterers is sufficiently large,
Assuming these elementary back-scatterers are large in number and independent in nature,  
  then the Central Limit Theorem is applicable \cite{Goodman_Springer_1975}.
Consequently the interference of these randomly phased responses can be considered as
  a random walk on the 2D complex plane \cite{Goodman_JOptSocAm_76}.  
Specifically the real part $A_r$ as well as the imaginary part $A_i$ of the observed SAR signal $A$ can be considered as
  random variables from uncorrelated Gaussian distributed stochastic processes with zero means and indentical variances $\sigma^2/2$  \cite{Lee_CRCPress_2009}. 
Their probability density functions (PDF) are given as:

@@

Systems were then engineered and commercial POLSAR satellites, e.g. RadarSat, TerraSAR, Alos-PalSAR were launched.
However due to operational restrictions on payload, power and data-downlink, the satellites also offer dual-channel polarimetric SAR modes as a compromise between theoretical polarimetric requirements and practical limitations.
Initially, to the end user, while dual-channel partial polarimetry may not provide the complete polarimetric signature of the target, the loss is compensated for by either better resolutions or higher coverage and lower cost.

@@

For a dihedral target, whose main axis is at an angle $\theta$ with respect to the horizontal direction, the Sinclair matrix is given as:
\begin{equation}
\Gamma_{dih} = 
\left(
\begin{array}{c c}
\cos(2\theta) & \sin(2\theta) \\
\sin(2\theta) & \cos(2\theta)
\end{array}
\right)
\end{equation}
The even bounce phenomena is detectable as its polarimetric response, which is then:
\begin{equation}
J_r = \left(
\begin{array}{c c}
\cos(2\theta) & \sin(2\theta) \\
\sin(2\theta) & \cos(2\theta)
\end{array}
\right) \cdot
\left(
\begin{array}{c}
1 \\
-i
\end{array}
\right) = \left(
\begin{array}{c}
1 \\
-i
\end{array}
\right) 
\end{equation}
In \cite{Raney_2007_IGARSS}, Raney shows that hybrid circular modes bring many conveniences to hardware calibration.
Also in a follow up study \cite{Raney_2007_TGRS}, he demonstrated the classification power of hybrid polarimetry.
Furthermore, Lardeux \cite{Lardeux_2007_ESASP} found that for target classification using SVM techniques, compact polarimetry mode shows comparable results to full polarimetric data.
Notably, the hybrid architectural concept has been adopted and brought towards implementation by NASA's upcoming Desdynl satellite \cite{Raney_2009_RadarConf}.

@@

%Clearly should such PDF being estimatable from the data it-self, that would potentially presents an good opportunity for this class of speckle filters.
Other distributions that have been proposed include the heavy tailed Rayleigh distribution \cite{Sun_2006_ICVES}, Weibull distribution \cite{Nezry_1997_IGARSS}, beta distribution \cite{Nezry_1997_IGARSS}, Rician distribution \cite{Lewinski_1983_TAntennaPropagation} \dots 
Evidently, the performance of these filters are dependent on the underlying distribution of the back-scattering coefficient on the particular surface being imaged. 
In this context, the community appears to focus on using Bayesian inference to choose the most suitable distributions \cite{Walessa_2000_TGRS}.
One of the additive and homoskedastic SAR model's benefits is the possible hypothesis that the underlying distribution of the back scattering coefficient is estimable discussed in subsequent chapters.
%In our approach, we hypothesize that the underlying distribution of the back scattering coefficient is estimable, at least in the log-transformed domain, as will be described later.

@@

Compared to SAR technologies, most POLSAR techniques are relatively recent. 
Published literatured on POLSAR speckle filtering indicate that filtering the covariance matrix off-diagonal elements is still an open problem.
The problem can be traced back to the lack of a statistical model for the covariance matrix in polarimetry.
%This has the consequences that very few methodologies to statistically simulate POLSAR is porposed in literature and validations for such proposals are pretty primatives.
%is the application of such framework is shown to be beneficial academically.

@@

Further attempts have been made towards statistically modelling the off-diagonal elements but the application of these results for speckle suppression of off-diagonal element have not been conclusive.
For instance Martinez et. al.\cite{Lopez_2008_TGRS} proposed a model based polarimetric filter.
They first proposed to model $S_hS_v^*$ as having both additive and multiplicative components:
\begin{equation}
S_pS_q^* = \psi |\rho | e^{i \theta} + \psi \bar{z}_n N_c (1-n_m) e^{i \theta} + \psi (n_{ar} + i n_{ai}) 
\end{equation}
where
	$\psi = \sqrt{ E \left( |S_p|^2 \right) E \left( |S_q|^2 \right) }$ represents the average joint power in the two channels, 
	$\rho = |\rho| e^{i \theta} = \frac{E \left( S_pS_q^* \right)}{ \sqrt{ E \left( |S_p|^2 \right) E \left( |S_q|^2 \right) } }$ is the complex correlation coefficient that characterises the correlation among the channels,
	$N_c = \frac{\pi}{4} |\rho| F_{1,2} ( 1/2,1/2,2,|\rho|^2 )$ basically contains the same information as the complex correlation coefficient.
	$F_{1,2} ( a,b,c,d )$ is the Gauss hyper-geometric function,
	$n_m$ is the first speckle noise component that is multiplicative with $E(n_m)=1$ and $var(n_m)=1$,
	$n_{ar} + i n_{ai}$ is the second additive speckle noise component with $E(n_{ar}) = E(n_{ai}) = 0$ and $var(n_{ar}) = var(n_{ai}) = \frac{1}{2} (1-|\rho|^2)^{1.32}$.

The model is overly complicated and to do speckle filtering, various approximations have to be used.

Another different approach is proposed bu Lee \cite{Lee_CRCPress_2009}  to model $S_hS_v^*$.
In this method, the phase difference is defined as:
\begin{equation}
\phi = \arg{ \left( S_pS_q^* \right) }
\end{equation}
and the PDF for the phase difference as:
\begin{equation}
pdf(\phi) = \frac{\Gamma(3/2) (1-|\rho|^2) \beta}{2 \sqrt{\pi} \Gamma(1) (1-\beta^2)} + \frac{(1-|\rho|^2)}{2 \pi} F_{1,2}(1,1,1/2,\beta^2)
\end{equation}
where
	$\beta = |\rho| cos(\phi-\theta)$ and
	$F_{1,2} ( a,b,c,d )$ is the Gauss hyper-geometric function.

@@

Because of the variaties of  speckle filtering available, it is hence crucial to have a framework for 
	(a) modelling and simulation, which provide the ground-truth data in different scenarios of underlying distribution, and
	(b) analysing the interdependency between noise and the underlying signal, as well as the dependence of each filter's performance on the different underlying distribution scenarios and
	(c) establishing consistent criteria to evaluate speckle suppression power as well as the adaptive capabilities in reconstructing the underlying signal.

>>>>

The stochastic nature of SAR data arises due to the process of SAR image formation. Specifically, since one SAR resolution cell is very big in comparison to the radar wavelength, it normally consists of many elementary scatterers. The data point of each SAR pixel then arises from the combined interference effect of the back-scattered waves from these different scatterers. This effect causes natural variation in SAR intensity image even when the underlying scene is highly homogeneous.

SAR speckle phenomena are explained as the interference of many coherent but dephased back-scattering components;
  each reflecting from different and distributed elementary scatterers \cite{Oliver_ProcIEEE_1963, Leith_ProcIEEE_1971}. 
The random nature of the SAR data arises
  due to the unknown and varying location, height and thus distance of each elementary scatterer;
  which, in turn, generates random phases in their back-scattering responses.
%Assuming the number of these elementary back scatterers is sufficiently large,
Assuming these elementary back-scatterers are large in number and independent in nature,  
  then the Central Limit Theorem is applicable \cite{Goodman_Springer_1975}.
Consequently the interference of these randomly phased responses can be considered as
  a random walk on the 2D complex plane \cite{Goodman_JOptSocAm_76}.  
Specifically the real part $A_r$ as well as the imaginary part $A_i$ of the observed SAR signal $A$ can be considered as
  random variables from uncorrelated Gaussian distributed stochastic processes with zero means and indentical variances $\sigma^2/2$  \cite{Lee_CRCPress_2009}. 
Their probability density functions (PDF) are given as:

@@

Systems were then engineered and commercial POLSAR satellites, e.g. RadarSat-2, TerraSAR, Alos-PalSAR were launched.
However due to operational restrictions on payload, power and data-downlink, the satellites also offer dual-channel polarimetric SAR modes as a compromise between theoretical polarimetric requirements and practical limitations.
Initially, to the end user, while dual-channel partial polarimetry may not provide the complete polarimetric signature of the target, the loss is compensated for by either better resolutions or higher coverage and lower cost.

@@

For a dihedral target, whose main axis is at an angle $\theta$ with respect to the horizontal direction, the Sinclair matrix is given as:
\begin{equation}
\Gamma_{dih} = 
\left(
\begin{array}{c c}
\cos(2\theta) & \sin(2\theta) \\
\sin(2\theta) & \cos(2\theta)
\end{array}
\right)
\end{equation}
The even bounce phenomena is detectable as its polarimetric response, which is then:
\begin{equation}
J_r = \left(
\begin{array}{c c}
\cos(2\theta) & \sin(2\theta) \\
\sin(2\theta) & \cos(2\theta)
\end{array}
\right) \cdot
\left(
\begin{array}{c}
1 \\
-i
\end{array}
\right) = \left(
\begin{array}{c}
1 \\
-i
\end{array}
\right) 
\end{equation}
In \cite{Raney_2007_IGARSS}, Raney shows that hybrid circular modes bring many conveniences to hardware calibration.
Also in a follow up study \cite{Raney_2007_TGRS}, he demonstrated the classification power of hybrid polarimetry.
Furthermore, Lardeux \cite{Lardeux_2007_ESASP} found that for target classification using Support Vector Machine (SVM) techniques, compact polarimetry mode shows comparable results to full polarimetric data.
Notably, the hybrid architectural concept has been adopted and brought towards implementation by NASA's upcoming Desdynl satellite \cite{Raney_2009_RadarConf}.

@@

%Clearly should such PDF being estimatable from the data it-self, that would potentially presents an good opportunity for this class of speckle filters.
Other distributions that have been proposed include the heavy tailed Rayleigh distribution \cite{Sun_2006_ICVES}, Weibull distribution \cite{Nezry_1997_IGARSS}, beta distribution \cite{Nezry_1997_IGARSS}, Rician distribution \cite{Lewinski_1983_TAntennaPropagation}. 
Evidently, the performance of these filters are dependent on the underlying distribution of the back-scattering coefficient on the particular surface being imaged. 
In this context, the community appears to focus on using Bayesian inference to choose the most suitable distributions \cite{Walessa_2000_TGRS}.
One of the additive and homoskedastic SAR model's benefits is the possible hypothesis that the underlying distribution of the back-scattering coefficient is estimable discussed in subsequent chapters.
%In our approach, we hypothesize that the underlying distribution of the back scattering coefficient is estimable, at least in the log-transformed domain, as will be described later.

@@

Compared to SAR technologies, most POLSAR techniques are relatively recent. 
Published literature on POLSAR speckle filtering indicates that filtering the covariance matrix off-diagonal elements is still an open problem.
The problem can be traced back to the lack of a statistical model for the covariance matrix in polarimetry.
%This has the consequences that very few methodologies to statistically simulate POLSAR is porposed in literature and validations for such proposals are pretty primatives.
%is the application of such framework is shown to be beneficial academically.

@@

Further attempts have been made towards statistically modelling the off-diagonal elements but the application of these results for speckle suppression of off-diagonal element have not been conclusive.
For instance Martinez et. al.\cite{Lopez_2008_TGRS} proposed a model based polarimetric filter.
They first proposed to model $S_hS_v^*$ as having both additive and multiplicative components:
\begin{equation}
S_pS_q^* = \psi |\rho | e^{i \theta} + \psi \bar{z}_n N_c (1-n_m) e^{i \theta} + \psi (n_{ar} + i n_{ai}) 
\end{equation}
where
	$\psi = \sqrt{ E \left( |S_p|^2 \right) E \left( |S_q|^2 \right) }$ represents the average joint power in the two channels, 
	$\rho = |\rho| e^{i \theta} = \frac{E \left( S_pS_q^* \right)}{ \sqrt{ E \left( |S_p|^2 \right) E \left( |S_q|^2 \right) } }$ is the complex correlation coefficient that characterises the correlation among the channels,
	$N_c = \frac{\pi}{4} |\rho| F_{1,2} ( 1/2,1/2,2,|\rho|^2 )$ basically contains the same information as the complex correlation coefficient.
	$F_{1,2} ( a,b,c,d )$ is the Gauss hyper-geometric function,
	$n_m$ is the first speckle noise component that is multiplicative with $E(n_m)=1$ and $var(n_m)=1$,
	$n_{ar} + i n_{ai}$ is the second additive speckle noise component with $E(n_{ar}) = E(n_{ai}) = 0$ and $var(n_{ar}) = var(n_{ai}) = \frac{1}{2} (1-|\rho|^2)^{1.32}$.
The model is overly complicated and to do speckle filtering, various approximations have to be used.

Another different approach is proposed by Lee \cite{Lee_CRCPress_2009}  to model $S_hS_v^*$.
In this method, the phase difference is defined as:
\begin{equation}
\phi = \arg{ \left( S_pS_q^* \right) }
\end{equation}
and the PDF for the phase difference as:
\begin{equation}
pdf(\phi) = \frac{\Gamma(3/2) (1-|\rho|^2) \beta}{2 \sqrt{\pi} \Gamma(1) (1-\beta^2)} + \frac{(1-|\rho|^2)}{2 \pi} F_{1,2}(1,1,1/2,\beta^2)
\end{equation}
where
	$\beta = |\rho| cos(\phi-\theta)$ and
	$F_{1,2} ( a,b,c,d )$ is the Gauss hyper-geometric function.

@@

Because of the variaties of  speckle filtering available, it is hence crucial to have a framework for 
	(a) modelling and simulation, which provide the ground-truth data in different scenarios of underlying distribution, and
	(b) analysing the inter-dependency between noise and the underlying signal, as well as the dependency of each filter's performance on the different underlying distribution scenarios and
	(c) establishing consistent criteria to evaluate speckle suppression power as well as the adaptive capabilities in reconstructing the underlying signal.
